Self Regulation Ontology Retest Analyses

### 2018-10-16 code/SRO_Retest_Analyses.Rmd
removed exploratory analyses on systematic effects of time

## Systematic effects between time points

### Anova time effects

The type of ICC we have chosen does not take within subject (between session/systematic) variance in to account. This is why Weir recommends checking whether there is a significant change based on time and examining the SEMs. These systematic effects could be meaningful and important to account for for some measures (e.g. task measures that show learning effects).

Had we chosen another kind of ICC taking this source of variance into account (e.g. 2,1 or 2,k) they could have suggested that tasks have lower reliability.

Doing a simple t-test on the difference score alone would not be a very rigourous way of testing whether any change is meaningful because two distributions from both time points with error would be compared to each other. Fortunately there are ways to take the error for both measurements in to account.

To check whether a measure shows systematic differences between the two time points in a meaningful number of bootstrapped samples we can: check if the effect of time is significant in each bootstrapped sample and filter variables that have more than 5% of the boostrapped samples showing significant time effects.

Another way might be to compute confidence intervals using SEMs as described in the second half of Weir (2005) and check what percent of participants have scores that fall out of this range.

First we ask: Which variables have significant time effects in more than 5% of the bootstrapped samples?

23/74 survey measures
133/372 task measures

```{r}
boot_df %>%
  select(dv, p_time, task) %>%
  mutate(time_effect_sig = ifelse(p_time<0.05,1,0)) %>%
  group_by(dv)%>%
  summarise(pct_sig_time_effect = sum(time_effect_sig)/10,
            task = unique(task))%>%
  filter(pct_sig_time_effect>5) %>%
  arrange(task,-pct_sig_time_effect) %>%
  ungroup()%>%
  group_by(task) %>%
  summarise(count=n()) %>%
  mutate(total_num_vars = c(74, 372),
         prop = count/total_num_vars) %>%
  ggplot(aes(task, prop, fill=task))+
  geom_bar(stat="identity", alpha=0.5, width=0.5)+
  theme(legend.position = "none",
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(size=14))+
  ylab("Proportiono of bootstrap samples")+
  xlab("")+
  ylim(0,1)

ggsave('PropTimeEffects.jpg', device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures/", width = 3, height = 5, units = "in", limitsize = FALSE, dpi = 300)

```

### SEM CI calculation

Step 1: T = Grand mean + ICC * (Subject score - Grand mean)
Step 2: SEP = SD of both measurements * sqrt(1-ICC^2)

Here I calculate the proportion of subjects that move more than one standard error of prediction (SEP) in t2 compared to t1 for each measure.

This is odd to think about because the larger the ICC of a measure the smaller the SEP. So very small differences between the two time points can be categorized as 'meaningful' based on the tiny SEP.

What you are interested in is not necessarily whether individuals change at all between the two time points though. You want to know if this change is systematic in one direction.

Here I calculate the proportion of people showing 'meaningful' changes in one direction or the other. To integrate both of these direction I subtracted one from the other and filtered the variables that have more than 5% of the participants showing meaningful change in one direction over the other (so if a variable has 10 participants showing difference in one and 10 in the other direction this would cancel out but a variables with 15 people showing a positive and 5 negative change would remain).

```{r}
get_ind_ci = function(dv_var){
  matched = match_t1_t2(dv_var)
  grand_mean = mean(matched$score)
  grand_sd = sd(matched$score)
  dv_icc = get_icc(dv_var)
  sep = grand_sd * sqrt(1-(dv_icc^2))
  matched = matched %>%
    spread(time, score) %>%
    rename("t1"="1", "t2"="2") %>%
    mutate(true_score = grand_mean+(dv_icc*(t1-grand_mean)),
           ci_up = true_score+(1.96*sep),
           ci_low = true_score-(1.96*sep),
           t2_above_ci = ifelse(t2>ci_up,1,0),
           t2_below_ci = ifelse(t2<ci_low,1, 0))
    return(matched)
}

get_prop_out_ci = function(dv_var){
  get_ind_ci(dv_var) %>%
    summarise(prop_above_ci = sum(t2_above_ci)/n(),
              prop_below_ci = sum(t2_below_ci/n()))
}

#Create df of point estimate reliabilities
ind_ci_df <- data.frame(prop_above_ci = rep(NA, length(numeric_cols)),
                        prop_below_ci = rep(NA, length(numeric_cols)))

row.names(ind_ci_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
    ind_ci_df[numeric_cols[i], 'prop_above_ci'] <- get_prop_out_ci(numeric_cols[i])$prop_above_ci
    ind_ci_df[numeric_cols[i], 'prop_below_ci'] <- get_prop_out_ci(numeric_cols[i])$prop_below_ci
}

ind_ci_df$dv = row.names(ind_ci_df)
row.names(ind_ci_df) = seq(1:nrow(ind_ci_df))
ind_ci_df$task = 'task'
ind_ci_df[grep('survey', ind_ci_df$dv), 'task'] = 'survey'
ind_ci_df[grep('holt', ind_ci_df$dv), 'task'] = "task"
ind_ci_df = ind_ci_df %>%
  select(dv, task, prop_above_ci, prop_below_ci)

mean_diff_df = ind_ci_df %>%
  mutate(prop_one_direction = prop_above_ci-prop_below_ci) %>%
  filter(abs(prop_one_direction)>0.05) %>%
  arrange(-abs(prop_one_direction))

mean_diff_df
```

This doesn't make it easier to interpret whether it is a performance improvement or decline. In fact 'performance' isn't even necessarily the correct term here.

## Task Reliabilities

>>>>

knitr::include_graphics(paste0(fig_path, fig_name))
```
## Task Reliabilities
